\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}

\title{A Theoretical and Computational Framework for Simulating Limit Order Book Dynamics and Optimal Execution}
\author{AI Quant Finance Researcher}
\date{\today}

\begin{document}

\maketitle

\section*{Part A: Theoretical Framework}

\section{Formalisms of the Limit Order Book}

\subsection{Definitions}

\begin{itemize}
    \item \textbf{Order:} A tuple $O = (\epsilon_o, p_o, v_o, \tau_o)$, where $\epsilon_o \in \{+1, -1\}$ (buy/sell), $p_o \in \mathbb{R}^+$ is the price, $v_o \in \mathbb{Z}^+$ is the volume, and $\tau_o \in \mathbb{R}^+$ is the timestamp.
    \item \textbf{Limit Order:} An order to buy/sell a quantity at a specified price or better. It rests in the LOB until executed or cancelled.
    \item \textbf{Market Order:} An order to buy/sell a quantity immediately at the best available price(s).
    \item \textbf{Tick Size ($\theta$):} The minimum price increment allowed by the exchange.
    \item \textbf{Queue Position:} An order's rank in the FIFO queue at a given price level.
    \item \textbf{Liquidity:} Characterized by tightness (spread), depth (volume at price levels), and resiliency (recovery speed).
    \item \textbf{Mid-Price:} $M(t) = \frac{P_a(t) + P_b(t)}{2}$, where $P_a(t)$ is the best ask and $P_b(t)$ is the best bid.
    \item \textbf{Microstructure Noise:} Deviations of the observed price from the efficient price due to discrete trading mechanics.
\end{itemize}

\section{Stochastic Modeling of LOB Event Flows}

\subsection{Order Arrivals}

\begin{itemize}
    \item \textbf{Poisson Process:} Constant intensity $\lambda(t) = \lambda$. Inter-arrival times are exponentially distributed.  $P(N(t) = k) = \frac{(\lambda t)^k e^{-\lambda t}}{k!}$.
    \item \textbf{Hawkes Process:} Self-exciting process. Intensity $\lambda(t) = \mu + \sum_{t_i < t} g(t - t_i)$. Exponential kernel: $g(t) = \alpha e^{-\beta t}$. Stationarity condition: $\frac{\alpha}{\beta} < 1$.
\end{itemize}

\subsection{Cancellations}

Modeled by a hazard rate $\lambda_c(t, O, LOB_t)$ dependent on order characteristics and market state.

\subsection{Price Impact}

\begin{itemize}
    \item \textbf{Linear:} $I(Q) \propto Q$ (small trades).
    \item \textbf{Square-root:} $I(Q) \propto \sqrt{Q}$ (large trades).
\end{itemize}

\section{Optimal Execution: Almgren-Chriss}

\subsection{Problem Statement}

Minimize expected cost $E[C]$ plus variance penalty $\lambda V[C]$, where $C$ is the implementation shortfall and $\lambda$ is the risk aversion.

\subsection{Market Impact Model}

\begin{itemize}
    \item \textbf{Permanent:} $\gamma v_k \tau$.
    \item \textbf{Temporary:} $\eta v_k$.
\end{itemize}

\subsection{Optimal Trajectory}

$x(t) = X_0 \frac{\sinh(\kappa(T-t))}{\sinh(\kappa T)}$, where $\kappa = \sqrt{\frac{\lambda \sigma^2}{\eta}}$.

\section{Simulator Design and Validation}

\subsection{Deliverables}

\begin{itemize}
    \item \texttt{LimitOrderBook} class.
    \item Event generators (Poisson, Hawkes).
    \item Simulation engine.
    \item Metrics and outputs (mid-price, slippage).
\end{itemize}

\subsection{Unit Tests}

\begin{enumerate}
    \item \textbf{Volume Conservation:} Verify total traded volume equals market order size.
    \item \textbf{Price-Time Priority:}  Confirm execution against best price levels first, then FIFO within each level.
    \item \textbf{Impact Scaling:} Check if slippage scales linearly for small trades.
    \item \textbf{Hawkes Clustering:} Ensure Hawkes process generates clustered events (higher variance of inter-arrival times than Poisson).
\end{enumerate}

\end{document}